The CrypTool Script: \\ Cryptography, Mathematics and More \\ \vspace{30pt} {\large Background reading\\%*[-8pt] for the free e-learning program CrypTool\\ (10th edition -- distributed with CrypTool version 1.4.30)} \vspace{60pt}

### 2009-11-27 script/en/chapters/what-is-sage.tex
Script[de+en]: Many small changes, e.g. Sage samples collected in files, use of iload; described how to deal with Sage notebook using interact; substituted href by url.

%% \centering statt \begin{center} ... \end{center}, da dies zusätzlich vertical space erzeugt.
\noindent In the meantime there are lots of PDF or HTML documents about using Sage,
so we name only a few of them: A good starting point is the ``Library''
from the Sage main page, the ``Documentation Project'' and the teaching
material\footnote{%
- ``Library'': {\centering \url{http://www.sagemath.org/library/index.html}},\\
- ``Documentation Project'': {\centering \url{http://wiki.sagemath.org/DocumentationProject}},\\
- ``Teaching'': {\centering \url{http://wiki.sagemath.org/Teaching_with_SAGE}}.
}.

@@

% ---------------------------------------------------------------------------
\section*{Sage user interfaces}
Sage is available free of charge and can be downloaded from the
following website:
\begin{center}
  \url{http://www.sagemath.org} \\
\end{center}
The default interface to Sage is command line based, as shown in
figure~\ref{fig:sage_cmd_interfaces}. However, there is a
graphical user interface to the software as well in the form of the
Sage notebook (see figure~\ref{fig:sage_gui_interfaces}). We can even
use Sage online at different servers, e.g.
\begin{center}
\url{http://www.sagenb.org} or\\
\url{http://sage.mathematik.uni-siegen.de:8000}
\end{center}
without having to install Sage locally.

@@

\footnotetext{%
To start the Sage gui: Enter notebook() at the Sage prompt, and then your favorite
browser (Iceweasel, Firefox, IE, ...) is started e.g. with the URL \url{http://localhost:8000}.
}

@@

\noindent and you can then proceed to use on the command line anything that is
defined in that Sage script\footnote{%
Notes:

@@

\noindent\hangindent=6pt\makebox[6pt][l]{-}If you run your script as above,
then Sage first parses your script,
writes it to another file called ``primroots.py'' (note the ``.py''
extension), adds all necessary variables to "primroots.py" as well as
writing any import statements to that file. That way, your Sage script
is executed as if you had typed the definitions in your script to the
Sage command line. A difference may be that output needs a print statement.

@@

Normally we also want to edit our own Sage script and reload the
content of the changed script into Sage again. In that case, you can
use the command \verb!attach()!. This looks in our Sage session as
follows (you also can apply \verb!attach()! directly after loading the script,
even before having changed the script):

@@

\vspace{30pt}
\noindent Some more hints:
\begin{itemize}
  \item To get the version of your Sage environment: \texttt{version()}
  \item To move quickly to the Sage code examples in this script,
    \begin{itemize}
      \item either look in the index at \verb#Sage -> Code examples#,
      \item or have a look at the appendix ``\nameref{sc:List-of-Sage-Code-Examples}''.
  \end{itemize}
\end{itemize}

>>>>

%% \centering statt \begin{center} ... \end{center}, da dies zusätzlich vertical space erzeugt.
\noindent In the meantime there are lots of PDF and HTML documents about using Sage,
so we name only a few of them as a good starting point\footnote{%
- ``Library'': {\centering \url{http://www.sagemath.org/library/index.html}},\\
- ``Documentation Project'': {\centering \url{http://wiki.sagemath.org/DocumentationProject}},\\
- ``Teaching'': {\centering \url{http://wiki.sagemath.org/Teaching_with_SAGE}}.
}.

@@

% ---------------------------------------------------------------------------
\section*{Sage user interfaces}
Sage is available free of charge and can be downloaded from the
following website:
\begin{center}
  \url{http://www.sagemath.org} \\
\end{center}
The default interface to Sage is command line based, as shown in
figure~\ref{fig:sage_cmd_interfaces}. However, there is a
graphical user interface to the software as well in the form of the
Sage notebook (see figure~\ref{fig:sage_gui_interfaces}). We can even
use Sage notebooks\footnote{%
Further details about Sage notebooks can be found at 
chapter~\ref{ec:Sage_Massierer} 
(``\nameref{ec:Implementing-for-Education}''
$\Rightarrow$ ``\nameref{ec:Sage_Massierer}'')
                      }
online at different servers, without having to install Sage locally, e.g:
\begin{center}
\url{http://www.sagenb.org} or\\
\url{http://sage.mathematik.uni-siegen.de:8000}
\end{center}
.

@@

\footnotetext{%
To start the Sage gui locally: Enter notebook() at the Sage prompt, and then your favorite
browser (Iceweasel, Firefox, IE, ...) is started e.g. with the URL \url{http://localhost:8000}.
}

@@

\noindent and you can then proceed to use on the command line any variable or function
defined in that Sage script\footnote{%
Notes:

@@

\noindent\hangindent=6pt\makebox[6pt][l]{-}If you run your script as above,
then Sage first parses your script,
writes it to another file called ``primroots.py'' (note the ``.py''
extension), adds all necessary variables to "primroots.py" as well as
writing any import statements to that file. That way, your Sage script
is executed as if you had typed the definitions in your script to the
Sage command line. An important difference isthat all output needs a
\verb!print! statement.

@@

Normally we also want to edit our own Sage script and reload the
content of the changed script into Sage again. In that case, you can
use the command \verb!attach()! (you also can apply \verb!attach()!
directly after loading the script, even before having changed the script;
and you can even omit \verb!load()!, as this is contained in \verb!attach()!):

@@

\vspace{20pt}
If you prefer to see the output of an attached file as if you would have
typed in the commands on the commandline directly (not only what is shown
via \verb!print!) then you could use the command \verb!iload()!:
Each line is loaded one at a time. To load the next line, you have to
press the \verb!Enter! key. You have to repeatedly press the \verb!Enter! key
until all lines of the Sage script are loaded into the Sage session.

\texttt{sage: iload primroots.sage}

% \vspace{30pt}
\newpage
\noindent Some more hints:
\begin{itemize}
  \item To get the version of your Sage environment: \texttt{version()}
  \item To move quickly to the Sage code examples in this script,
    \begin{itemize}
      \item either look in the index at \verb#Sage -> Code examples#,
      \item or have a look at the appendix ``\nameref{sc:List-of-Sage-Code-Examples}''.
    \end{itemize}
  \item The source code of the Sage samples in this script is delivered as
        Sage program files within the CrypTool setup program.
        After installing CrypTool you find them within the subdirectory \verb#sage#
        within the CrypTool directory:\\
        - SAGE-Samples-in-Chap01.sage \\
        - SAGE-Samples-in-Chap02.sage \\
        - SAGE-Samples-in-Chap03.sage \\
        - SAGE-Samples-in-Chap04.sage
\end{itemize}

### 2009-11-27 script/en/script-en.tex
fixed math mode in headings resulting in broken PDF bookmarks with \texorpdfstring; fixed float warnings by replacing placement [h] with [ht]; fixed warning about PDF version 1.5 by changing it to 1.4 in coverv1_*.pdf
enable verbose in footnotes using \VerbatimFootnotes; fixed some TODOs
CT script: Sage parts finished in (en); translation to German almost finished
CT script: Minor corrections, mostly layout, de + en; final now for CT 1.4.30 Beta04
document used sw on page 2
Script[de+en]: hanging indent in footnotes; corrected links to old appendix parts
added contents overview based on shortt

\usepackage[flushmargin]{footmisc}
\usepackage{fullpage}
\usepackage[hyphens]{url} % options "hyphens" enables linebreaks in \url{}
% Literaturverzeichnis kapitelweise
\usepackage[labelstoglobalaux]{bibunits} % ohne labelstoglobalaux scheint man thebibilograpy und bibtex nicht mischen zu können
%\bibliographyunit[\section] % bibunits für das ganze Dokument aktivieren
\defaultbibliography{references}
\defaultbibliographystyle{abbrv}
\usepackage{float}
\usepackage[colorlinks=true,linkcolor=black,citecolor=black,urlcolor=black,bookmarksnumbered=true,pdfpagelabels,plainpages=false,hyperfootnotes=false]{hyperref}
%To reduce the number of hyperlink related 'pdfTeX warning's we use options 
%pdfpagelabels, plainpages=false and, hyperfootnotes=false and place hyperref after float.
%Option debug helps finding problems.

\floatstyle{ruled}
\newfloat{sagecode}{!h}{loc}[chapter]
\floatname{sagecode}{Sage sample}
\newfloat{cryptoprocedure}{!h}{lop}[chapter]
\floatname{cryptoprocedure}{Crypto Procedure}

@@

\vspace{500pt}
Source cover photograph: \url{www.photocase.com}, Andre Guenther

@@

\newpage
\pdfbookmark[0]{\contentsname}{Contents}
\tableofcontents

@@

\clearpage\phantomsection
\addcontentsline{toc}{chapter}{List of Sage Code Examples}
\listof{sagecode}{List of Sage Code Examples}

>>>>

\usepackage[flushmargin]{footmisc}
\usepackage{fullpage}
\usepackage[hyphens]{url} % options "hyphens" enables linebreaks in \url{}
% Literaturverzeichnis kapitelweise
\usepackage[labelstoglobalaux]{bibunits} % ohne labelstoglobalaux scheint man thebibilograpy und bibtex nicht mischen zu können
%\bibliographyunit[\section] % bibunits für das ganze Dokument aktivieren
\defaultbibliography{references}
\defaultbibliographystyle{abbrv}
\usepackage{float}
\usepackage{shorttoc}
\usepackage[colorlinks=true,linkcolor=black,citecolor=black,urlcolor=black,bookmarksnumbered=true,pdfpagelabels,plainpages=false,hyperfootnotes=false]{hyperref}
%To reduce the number of hyperlink related 'pdfTeX warning's we use options 
%pdfpagelabels, plainpages=false and, hyperfootnotes=false and place hyperref after float.
%Option debug helps finding problems.

\floatstyle{ruled}
\newfloat{sagecode}{!ht}{loc}[chapter]
\floatname{sagecode}{Sage sample}
\newfloat{cryptoprocedure}{!ht}{lop}[chapter]
\floatname{cryptoprocedure}{Crypto Procedure}

@@

\vspace{460pt}
Source cover photograph: \url{www.photocase.com}, Andre Guenther\\

Typesetting software: \LaTeX

Version control software: Subversion

@@

\newpage
\pdfbookmark[0]{Contents Overview}{ShortContents}
\shorttoc{Contents Overview}{0}
\clearpage\phantomsection
\pdfbookmark[0]{\contentsname}{Contents}
\tableofcontents

@@

\clearpage\phantomsection
\addcontentsline{toc}{chapter}{List of Sage Code Examples}
\listof{sagecode}{List of Sage Code Examples}
\label{sc:List-of-Sage-Code-Examples}

\vspace{30pt}
\begin{itemize}
  \item All samples have been tested with Sage v 4.2.1 (Release Date 2009-11-14).
  \item The source code of the Sage samples in this script is delivered as
        Sage program files within the CrypTool setup program.
        After installing CrypTool you find them within the subdirectory \verb#sage#
        within the CrypTool directory:\\
        - SAGE-Samples-in-Chap01.sage \\
        - SAGE-Samples-in-Chap02.sage \\
        - SAGE-Samples-in-Chap03.sage \\
        - SAGE-Samples-in-Chap04.sage
\end{itemize}
